Implicit yaddayadda

### 2011-10-11 papers/PLDI2012/implicitization.tex
+blurb SMT

\newcommand{\abstr}[1]{#1^\sharp}
\newcommand{\parts}[1]{\mathscr{P}(#1)}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\widening}{\mathop{\triangledown}}

@@

\subsection{SMT-solving}
TODO

>>>>

\newcommand{\abstr}[1]{#1^\sharp}
\newcommand{\parts}[1]{\mathscr{P}(#1)}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\widening}{\mathop{\triangledown}}

@@

\subsection{SMT-solving}
Boolean satisfiability (SAT) is the canonical NP-complete problem: given a propositional formula (e.g. $(a \lor \neg b) \land (\neg a \lor b \lor \neg c)$), decide whether it is satisfiable --- and, if so, output a satisfying assignment.
Despite an exponential worst-case complexity, the DPLL algorithm \cite{Kroening_Strichman_08,Handbook_SAT} solves many useful SAT problems in practice.

SAT was extended to \emph{satisfiability modulo theory} (SMT): in addition to propositional literals, SMT formulas admit atoms from a theory.
For instance, the theories of linear integer arithmetic (LIA) and linear real arithmetic (LRA) have atoms of the form $a_1 x_1 + \dots + a_n x_n \bowtie C$ where $a_1,\dots,a_n,C$ are integer constants, $x_1\,dots,x_n$ are variables (interpreted over $\ZZ$ for LIA and $\RR$ or $\QQ$ for LRA), and $\bowtie$ is a comparison operator $=,\neq,<,\leq,>,\geq$.
Satisfiability for LIA and LRA is NP-complete, yet tools based on DPLL(T) approach \cite{Kroening_Strichman_08,Handbook_SAT} solve many useful SMT problems in practice. All these tools provide a \emph{satisfying assignment} if the problem is satisfiable.

Most SMT solvers, including Z3%
\footnote{\url{http://research.microsoft.com/en-us/um/redmond/projects/z3/}}
Yices,%
\footnote{\url{http://yices.csl.sri.com/}}
and all those supporting the full SMTLIB2 standard \cite{BarST-SMTLIB},
offer an \emph{incremental} interface: the client program specifies the formula as an initially empty conjunction, to which additional constraints are added, and calls a ``check'' function answering whether it is satisfiable; 
it may then backtrack some of the constraints and add other ones without restarting from scratch.
